\subsection{M1: Check Null Mutator}
M1, the Check Null Mutator, inserts null checks on objects prior to field access instructions.
This has practical uses in generating fixes for unexpected \java{NullPointerException}s.
The two opcodes that indicate a field access are \jvm{PUTFIELD} and \jvm{GETFIELD} which write to and read from an object's field, respectively.
M1 executes those opcodes only if the object whose field they are accessing is not null.

\begin{Java}[%
    linebackgroundcolor={%
      \ifnum 0<\value{lstnumber} \color{white}\fi%
      \ifnum 4=\value{lstnumber} \color{diff add}\fi%
      % \ifnum 3<\value{lstnumber} \color{red}\fi,%
    },%
    caption={Sample Java With Simulated M1},%
    label={lst:m1:java},%
  ]
  Rct r = new Rct(1, 2);
  int width;

  if (null != r)
      width = r.width;
\end{Java}

\begin{Java}[%
    linebackgroundcolor={%
      \ifnum 0<\value{lstnumber} \color{white}\fi%
      \ifnum 6=\value{lstnumber} \color{diff add}\fi%
      \ifnum 7=\value{lstnumber} \color{diff add}\fi%
      \ifnum 8=\value{lstnumber} \color{diff add}\fi%
      \ifnum 9=\value{lstnumber} \color{diff add}\fi%
      \ifnum 10=\value{lstnumber} \color{diff add}\fi%
      \ifnum 14=\value{lstnumber} \color{diff add}\fi%
      \ifnum 15=\value{lstnumber} \color{diff add}\fi%
    },%
    caption={ASMified Version of Listing~\ref{lst:m1:java}},%
    label={lst:m1:jvm},%
  ]
  visitTypeInsn(NEW, "Rct");
  visitInsn(DUP);
  visitInsn(ICONST_1);
  visitInsn(ICONST_2);
  visitMethodInsn(INVOKESPECIAL... "<init>"...);
  visitVarInsn(ASTORE, 1);
  visitInsn(ACONST_NULL);
  visitVarInsn(ALOAD, 1);
  el l0 = new Label();
  visitJumpInsn(IF_ACMPEQ, l0);
  visitVarInsn(ALOAD, 1);
  visitFieldInsn(GETFIELD, "Rct", "width", "I");
  visitVarInsn(ISTORE, 2);
  visitLabel(l0);
  visitFrame(Opcodes.F_APPEND,1, ...);
\end{Java}

Listing~\ref{lst:m1:java} shows an excerpt of a sample Java program.
\java{width} may be assumed to be a valid field in the \java{Rct} class.
If one assumes that the lines with no background color are part of the original source code, the addition of the line highlighted in green demonstrates the Java equivalent of M1's bytecode manipulation.

Listing~\ref{lst:m1:jvm} is an excerpt of an ASMified view of the class file compiled from Listing~\ref{lst:m1:java}'s source file.
Here, the green highlight indicates lines resulting from the addition of the M1-imitating \java{null} check in Listing~\ref{lst:m1:java}.
One interesting thing to note in this sample is the use of \jvm{ASTORE} and \jvm{ALOAD} to first store \java{r}'s object reference into a local variable (which pops the reference from the stack) and then load that same reference from that variable back onto the stack.
This seems to be wasted effort, as the result of \jvm{IF_ACMPEQ} does not change based on the order of its operands;
as such, our implementation of M1 does not use \jvm{ASTORE} or \jvm{ALOAD}, instead executing \jvm{ACONST_NULL} and \jvm{IF_ACMPEQ} in immediate succession.

The general algorithm for our implementation of M1 is:
\begin{enumerate}
\item
  Duplicate the objectref on top of the stack
\item
  Load a \java{null} reference onto the stack
\item
  Compare the references' equality
\end{enumerate}
\begin{minipage}[t]{0.45\linewidth}
  If the references are the same (\ie{} \java{null}):
  \begin{enumerate}[%
    leftmargin=*,%
    label={(\arabic*a)}
    ]
    \setcounter{enumi}{3}
  \item
    Skip the original opcode
  \item
    Pop the unneeded object reference from the stack
  \item\label{itm:m1:load default}
    If the original opcode was \jvm{GETFIELD}, load a default value onto the stack.
  \item
    Return to normal execution
  \end{enumerate}
\end{minipage}
\hfill\vline\hfill
\begin{minipage}[t]{0.5\linewidth}
  If the references are \textit{not} the same (\ie{} are not \java{null}):
  \begin{enumerate}[%
    leftmargin=*,%
    label={(\arabic*b)}
    ]
    \setcounter{enumi}{3}
  \item\label{itm:m1:execute original}
    Execute the original field access opcode
  \item\label{itm:m1:return}
    Return to normal execution
  \end{enumerate}
\end{minipage}

\vspace*{2ex}

M1's logic is implemented in \java{visitFieldInsn()}.
Two labels are used in the implementation.
The first is placed immediately prior to the opcodes implementing the (a) branch above.
This label is jumped to conditionally with \jvm{IF_ACMPEQ}, whose arguments are a duplicate of the original object reference and a \java{null} reference.
If that jump is not performed then the execution is allowed to fall through to \ref{itm:m1:execute original}.
A second label is placed immediately \textit{after} \ref{itm:m1:load default}.
This label is jumped to unconditionally with a \jvm{GOTO} in \ref{itm:m1:return}.

The branch in step~\ref{itm:m1:load default} occurs only in Java source of the M1 itself;
there is not corresponding jump in the bytecode.
The opcode used to load the default value onto the stack depends on the value of the \java{type} argument passed to \java{visitFieldInsn()}.
The relationships between \java{type} and those opcodes are illustrated in Table~\ref{tab:m1:default opcode}.

\begin{table}
  \centering
  \begin{tabular}{l l l}
    \toprule
    Value of \java{type} & Opcode & Loaded Value \\
    \midrule
    \java{"I"} & \jvm{ICONST_0} & \java{int} value 0 \\
    \java{"F"} & \jvm{FCONST_0} & \java{float} value 0 \\
    \java{"J"} & \jvm{LCONST_0} & \java{long} value 0 \\
    \java{"D"} & \jvm{DCONST_0} & \java{double} value 0 \\
    anything else & \jvm{ACONST_NULL} & \java{null} \\
    \bottomrule
  \end{tabular}
  \caption{Mapping \java{type} to Default Values}
  \label{tab:m1:default opcode}
\end{table}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../report"
%%% End:
